Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-07-21 Document/paper.tex
edit

@@

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.

The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback \citep{Debattista08, Bryan13, Butsky16,
  Chua19, Artale19}.  
The key concept unifying these results is that the baryon distribution
influences and correlates with the dark matter halo shape. 
Here we find that the broad tendency is that massive stellar discs
correlate with spherical dark matter distributions.

In order to explore this idea in the Auriga simulations we quantify
the correlation between halo shape with baryonic disc properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disc to total mass ratio and
the optical radius.

Figure \ref{fig:disc_correlations} shows the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
We find that the strongest correlations are found with the halo shape
measured at radii smaller than $0.12R_{200}\sim 28$kpc, which 
is close to the upper limit of the disc optical radius among our
simulation sample. 
The trend is such that halos with large triaxiality correlate with
high gas density and stellar discs with low mass and small size.
In turn, massive and large stellar discs within a low density gas
environment correlate with low halo triaxiliaty.

>>>>

\subsection{Correlation with baryonic disc properties}

The presence of baryons produce rounder dark matter halos than its DMO
counterpart.
To better understand this relationship we quantify the correlation
between halo shape with baryonic disc properties.  
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disc to total mass ratio and
the optical radius.

Figure \ref{fig:disc_correlations} shows the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
We find that the strongest correlations are found with the halo shape
measured at radii smaller than $0.12R_{200}$.
The  trend is such that halos with large triaxiality correlate with
high gas density and stellar discs with low mass and small size. 
In turn, massive and large stellar discs within with low density of
gas in its core correlate with low halo triaxiliaty.

@@

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.
The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback \citep{Debattista08, Bryan13, Butsky16,
  Chua19, Artale19}.  
The key concept unifying these results is that the baryon distribution
influences and correlates with the dark matter halo shape. 
Here we find that the broad tendency is that massive stellar discs
correlate with spherical dark matter distributions.
